---
title: "Endogeic Biomass"
author:
  - name: Quentin Schorpp
    email: qurntin.schorpp@ti.bund.de
date: "`r Sys.Date()`"
output: 
      html_document:
          toc: true
          toc_depth: 3
          number_sections: true
          fig_caption: true
          theme: cerulean
          highlight: haddock
          
---
```{r SetGlobalOptions, include=FALSE}
knitr::opts_chunk$set(fig.align="center",
                      message=FALSE,
                      error=FALSE,
                      warning=FALSE,
                      echo=FALSE)
knitr::opts_knit$set(root.dir = 'D://Quentin_Schorpp//Arbeitsprozess//git_repositories//F2_Earthworms')

options(scipen=1, digits=2)
set.seed(1234)
```


# Introduction  

I investigated how earthworm communities change in perennial energy farming systems. To assess the development in the long term, I chose an approach called Space-for-Time Substitution (Pickett, 1989). Instead of sampling sites over a long time period, I studied different aged sites. The sites were cropped either with the perennial bioenergy plant cup-plant (*Silphium perfoliatum*) or with annual Maize (*Zea mays*). The Main question of this investigation is wether older cup plant fields possess higher structural and functional biodiversity than younger ones and to compare cup-plant to maize. To distinguish the age-effect from environmental influences, several parameters regarding soil properties, climate and Management were collected.
Sampling took place 3-4 times in two consecutive years. In total 18 sites were surveyed, comprising 4 cup-plant age classes and maize. The sites were located at seven different locations, comprising fields from research and business. Each cup-plant age-class had three replicates, whereas maize plants were switched in the second year of sampling due to crop rotation and gained six replicates. To sample earthworms 4 pits with 0.25 m² were digged up to 20 cm in depth and both handsorting of soil and chemical extractioon with AITC took place. These 4 pits were randomly chosen at the field, hence they're nested in field.
9 earthworm species from 3 different ecological groups were identified and weighed.

Here I present the modeling of the biomass data of earthworms from the ecological group of endogeic earthworms. 

# Materials and methods  

I use R with RStudio to perform all analyses. For creating reports i use knitr with RStudio.

## Modeling procedure:

* I start with a global model, with a full set of parameters. 
  
    + The choice of parameters took place according to preknowledge  about the biology of earthworms. 
    + Management parameters, are not included up to now. 

* Since Abundance is count data, i choose poisson distribution in the first place, and switch to negative binomial if overdipersion can be detected.

* As random effects i chose field. I could have chosen sampling campaign, too and analyse the model with crossed random effects structure, but I'm also interested in change over the time of sampling.

* I included an offset "area", since i want to present the data per m² not per 0.25 m².

* several global models were formulated if there are different combinations of parameters, meaningful for analysis, but not able to analyse at once, i.e. because of  collinearity.

* Then I use Multimodel Inference/Averaging to find the model with the lowest AIC and determine it as the best model.

* Model validation is performed by 

    + plotting residuals against all covariates in the model as well as against all covariates not in the model
    + plotting a histogram of the residual
    + checking for overdispersion

* after having found a model with good model fit, i present the data for the means of environmental covariates and plot barplots for the age classes.

* errorbars are calculated, using the inverse link function with confidence intervals.

* Post-Hoc Multicomparisons are performed using Tukey Test with the glht() function.

# Analysis

```{r LoadData, results='hide', fig.show='hide'}
# Load Data
#setwd("D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/")
source("Data/GatherSource/Gather1.R") # Includes Correction for Watercontent
source("Data/GatherSource/Gather2.R")
source("Data/GatherSource/Gather3.R") # no adult endogeics separated
source("Data/GatherSource/CNRatio.R") # only for 1st order
```

```{r RequiredPackages}
source("Analysis/F2_EW_RequiredPackages.R")
source("Data/GatherSource/Functions/coldiss.R")
source("Data/GatherSource/Functions/panelutils.R")
source("Data/GatherSource/ggplotTheme.R")

lo <- par(mfrow=c(1,1))

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

```


### Data Pre-Processing  
#### Standardize Explanatory Variables  

```{r DataProcessing,}

## Create variable d(delta)samcam as the time difference of each sampling date to the first date, that each field was sampled; first date is Zero for all fields. ####
data$dsamcam <- with(data, round(c(rep(0,60),date[61:120]-date[1:60],date[121:180]-date[1:60]),0))

## Standardize all explanatory variables ####

# extract all numerical variables
std.var <- data[sapply(data,is.numeric)]
# exclude all response variables
std.var <- std.var[,-c(7:30,49:58)]

# standardize, each column is divided by their standard deviation, and the mean of each column is subtracted 
std.var <- as.data.frame(scale(std.var, center=TRUE, scale=TRUE))
colnames(std.var) <- paste("scl", colnames(std.var), sep = ".") # rename columns

data$area <- rep(0.25,180)

data <- cbind(data, std.var)


# Ordered or not??
# Since i found it too difficult to handle, i used non-ordered factors
data$age_class <- factor(data$age_class, ordered=TRUE)
data$samcam <- factor(data$samcam, ordered=FALSE)

#----
```


### Endogeic Biomass
**********************


#### Assess variability of random effects
```{r BoxPlots,echo=FALSE}
par(lo)

# Variability within sites
boxplot(endo.bm~age_class, data, col="grey", main="Variability within sites", xlab="sites orderd in decreasing age", ylab="anecic earthworm abundance")

# variability within sampling campaigns
boxplot(endo.bm~samcam, data,col="grey", main="Variability within sampling campaigns", xlab="seasons:\n autumn2012, spring2013, autumn2013", ylab="anecic earthworm abundance")

# variability within sites and sampling campaigns
boxplot(endo.bm~field.ID+samcam, data, las=2, col="grey", main="Variability within sites \n differing at sampling campaigns", xlab="field+seasons", ylab="anecic earthworm abundance")

dotchart(data$endo.bm, main="Cleveland dotplot for endogeic biomass")

```
 
 
#### xy plots for linear realtionships and interactions
```{r xyplots, eval=FALSE}
xyplot(endo.bm~pH|age_class,data=data)
xyplot(endo.bm~cn|age_class,data=data)
xyplot(endo.bm~clay|age_class,data=data)
xyplot(endo.bm~mc|age_class,data=data) 
xyplot(endo.bm~cn|pH,data=data) 
````


#### GLMM using glmmADMB


##### Model Formulation

Global Model and Multimodel Inference
```{r GLMM_anc, cache=TRUE, eval=FALSE, echo=TRUE}
## check data ####
str(data)

# Percentage of zero values in biomass data for anecic earthworms
nrow(data[data$endo.bm==0,])/nrow(data)

# Subset of all Biomass data not 0
datax <- subset(data, endo.bm>0)

# Outliers in CN ratio
which(data$cn>15)
data[which(data$cn>15),]$field

# Eliminate outliers
datay <- data[data$field!="EF_2011",]
#####



## Global Model Formulation ####
## Formulate several global models with regard to Collinearity in environmental variables, outliers and zero values
## [Note: First off, let's start with what a significant continuous by continuous interaction means. It means that the slope of one continuous variable on the response variable changes as the values on a second continuous change.]

# Gamma Distribution
####################

# without zero values // -> hurdle models??
## with soil moisture
endo.bm.glob1 <- glmmadmb(endo.bm ~ age_class*samcam + scl.mc + scl.pH*scl.cn + (1|field.ID) + offset(log(area)),data=datax,family="gamma")

# adding small values to Zero 
## With soil moisture
endo.bm.glob2 <- glmmadmb(I(endo.bm+0.0001) ~ age_class*samcam + scl.mc + I(scl.mc^2) + scl.pH*scl.cn + (1|field.ID) + offset(log(area)),data=data,family="gamma")

## With average Temperature and humidity
endo.bm.glob3 <- glmmadmb(I(endo.bm+0.0001) ~ age_class*samcam + I(scl.ats1^2) + scl.hum1 + scl.pH*scl.cn + (1|field.ID) + offset(log(area)),data=data,family="gamma")


# Normal distribution
#####################

## With soil moisture
## With total data
endo.bm.glob3 <- glmmadmb(endo.bm ~ age_class*samcam + scl.mc + scl.pH*scl.cn + (1|field.ID) + offset(log(area)),data=data,family="gaussian")
# without zero biomass 
endo.bm.glob4 <- glmmadmb(endo.bm ~ age_class*samcam + scl.mc + scl.pH*scl.cn + (1|field.ID) + offset(log(area)),data=datax,family="gaussian")
# without outliers for cn
endo.bm.glob5 <- glmmadmb(endo.bm ~ age_class*samcam + scl.mc + scl.pH*scl.cn + (1|field.ID) + offset(log(area)),data=datay,family="gaussian")

## With Average Temperature
## With total data
endo.bm.glob7 <- glmmadmb(endo.bm ~ age_class*samcam + I(scl.ats1^2) + scl.hum1 + scl.pH*scl.cn + (1|field.ID) + offset(log(area)),data=data,family="gaussian")

# offset is used due to personal advice by T. Onkelinx:
# "You better use an offset if you want to express the model in terms of m². Just add offset(log(0.25)) to the model."
#####

## Multimodel averaging ####
endo.bm.dredge1 <- dredge(endo.bm.glob1)
endo.bm.dredge2 <- dredge(endo.bm.glob2)
endo.bm.dredge3 <- dredge(endo.bm.glob3)
endo.bm.dredge4 <- dredge(endo.bm.glob4)
endo.bm.dredge5 <- dredge(endo.bm.glob5)
endo.bm.dredge6 <- dredge(endo.bm.glob6)
endo.bm.dredge7 <- dredge(endo.bm.glob7)
head(endo.bm.dredge1,10)
head(endo.bm.dredge2,10) # best AICc + total data
head(endo.bm.dredge3,10)
head(endo.bm.dredge4,10)
head(endo.bm.dredge5,10)
head(endo.bm.dredge6,10)
head(endo.bm.dredge7,10)
#####
````

Fit the best model
```{r, eval=-2, cache=TRUE, echo=TRUE}
# The best model with both glmer() and glmmadmb():
endo.bm.best <- glmer(I(endo.bm+0.0001) ~ age_class + scl.pH + (1|field.ID) + offset(log(area)), data=data, family=Gamma)

endo.bm.best <- glmmadmb(I(endo.bm+0.0001) ~ age_class + scl.pH + (1|field.ID) + offset(log(area)) ,data=data, family="gamma") 

# To exclude large fitted values: data[-c(69,70,71,72),]
````

Summary output
```{r Summary_endo.bm, echo=TRUE}
# Summary output
summary(endo.bm.best)

# ANOVA output
summary(aov(endo.bm.best))
```


##### Model Validation

```{r ModelValidation1_endo.bm, echo=TRUE}
# Confidence Intervals
confint(endo.bm.best)
coefplot2(endo.bm.best)

## Model validation ####
E1 <- resid(endo.bm.best, type="response")
F1 <- fitted(endo.bm.best, type="response")
P1 <- predict(endo.bm.best, type="response")

# Check Model outputs
par(mfrow=c(2,2),
    mar=c(4,4.5,1,2))
# Plot fitted vs. residuals
scatter.smooth(F1, E1, cex.lab = 1.5, ylab=" Residuals", xlab="Fitted values")
abline(h = 0, v=0, lty=2)
data$field[[69]]

# plot fitted vs. predicted
scatter.smooth(F1, P1, cex.lab = 1.5, ylab=" Residuals", xlab="Predicted values")
abline(h = 0, v=0, lty=2)

# Histogram of Residuals
hist(E1, prob=TRUE, main = "", breaks = 20, cex.lab = 1.5, xlab = "Response Residuals", col="PapayaWhip")
lines(density(E1), col="light blue", lwd=3)
lines(density(E1, adjust=2), lty="dotted", col="darkgreen", lwd=2) 

par(lo)

# negative fitted values:
which(F1<0)

## Residuals vs. covariates ####

# Dataset with all covariates IN the model
env1 <- cbind(E1, data[,c("endo.bm", "age_class", "pH")])
# covariates NOT in the model
env0 <- cbind(E1, data[,c(  "cn", "clay", "mc","samcam")])
env01 <- cbind(E1, data[,c(  "ats1", "ats2" ,"hum1", "dgO")])

# plot residual versus all covariates in the model 
par(mfrow=c(2,2),
     mar=c(3.8,4,1,3))
for (i in 2:length(env1)) {
  scatter.smooth(env1[,i],env1[,1], xlab=colnames(env1)[i], ylab="Residuals")
  abline(h=0, lty=2, col="red")
}

# plot residual versus all covariates NOT in the model
par(mfrow=c(2,2),
    mar=c(3.8,4,1,3))
for (i in 2:length(env0)) {
  scatter.smooth(env0[,i],env0[,1], xlab=colnames(env0)[i])
  abline(h=0, lty=2, col="red")
}

par(mfrow=c(2,2),
    mar=c(3.8,4,1,3))
for (i in 2:length(env01)) {
  scatter.smooth(env01[,i],env01[,1], xlab=colnames(env01)[i])
  abline(h=0, lty=2, col="red")
}
par(mfrow=c(1,1))
#####

## Some more plots ####
# Plots of Predicted Values
par(mfrow=c(2,2))
plot(data$age_class,P1, ylab="Predicted values")
plot(data$samcam,P1, ylab="Predicted values")
plot(data$pH,P1, ylab="Predicted values")



# Plots of fitted Values
par(mfrow=c(2,2))
plot(data$age_class,F1, ylab="Fitted values")
plot(data$samcam,F1, ylab="Fitted values")
plot(data$pH,F1, ylab="Fitted values")

par(lo)
#####
```


##### Post-Hoc Multicomparisons
```{r PostHocMulticomparisons_endo.bm, echo=TRUE}
# Pairwise comparisons
endo.bm.pairwise <- glht(endo.bm.best, mcp(age_class = "Tukey"))
endo.bm.pw.ci <- confint(endo.bm.pairwise)

# Confidence intervals including zero
endo.bm.pw.sig <- which(endo.bm.pw.ci$confint[,2]>0)
data.frame(names(endo.bm.pw.sig))


# Plot Errorbars
ggplot(endo.bm.pw.ci, aes(y = lhs, x = exp(estimate), xmin = exp(lwr), xmax = exp(upr))) + 
  geom_errorbarh() + 
  geom_point() + 
  geom_vline(xintercept = 1) +
  mytheme

# plot confidence intervals
par(mar=c(2,15,2,2))
plot(endo.bm.pw.ci) # only maize is significantly different from all SIlphie fields. Within SIlphie there are no differences

```


##### Extract Predictions and plot predictions with error bars
```{r ExtractPredicitons_endo.bm, echo=TRUE}
# create test data set with all covariates IN the model
# to predict for age_class only, take the mean of all continous covariates
endo.bm.td = expand.grid(age_class=unique(data$age_class),
                        scl.pH = mean(data$scl.pH),
                        area = 1)

# calculate confidence intervals for predictions from test dataset
endo.bm.pred <- cbind(endo.bm.td, predict(endo.bm.best, newdata = endo.bm.td, interval = "confidence")) 

# plot predictions with error bars // confidence intervals???
ggplot(endo.bm.pred, aes(x = age_class, y = exp(fit), ymin = exp(lwr), ymax = exp(upr))) + 
  geom_bar(stat="identity",position = position_dodge(1), col="454545", fill="grey") +
  geom_errorbar(position = position_dodge(1),col="black",width=0.15, size=0.15) + 
  geom_hline(xintercept = 1) +
  ylab("Biomass [g]") +
  xlab("Age Class") +
  scale_x_discrete(labels=c("Cm", "Sp_Y", "Sp_I1", "Sp_I2", "Sp_O")) +
  mytheme
````


Coefficients and Statistics
```{r, echo=TRUE}
# Claculate a whole lot of coefficients and statistics
endo.bm.pred <- within(endo.bm.pred, {
  AIC <- AIC(endo.bm.best)
  Rrandom <- summary(lm(fitted(endo.bm.best)~ data$endo.bm))$adj.r.squared
  Rsquared <- summary(lm(predict(endo.bm.best, type="response")~ data$endo.bm))$adj.r.squared  
})
head(endo.bm.pred)

endo.bm.stat = round(summary(endo.bm.best)$coefficients[, c(3,4)],4)
endo.bm.coef = coeftab(endo.bm.best)

endo.bm.env <- glmmadmb(endo.bm ~ scl.pH + (1|field.ID) + offset(log(area)),data=data,family="poisson")
endo.bm.pvalue <- anova(endo.bm.env, endo.bm.best)$"Pr(>Chi)"

endo.bm.OUT1 = endo.bm.pred
endo.bm.OUT2 = cbind(endo.bm.stat,endo.bm.coef,LogLikP=endo.bm.pvalue[2],fixef=fixef(endo.bm.best))
endo.bm.OUT1
endo.bm.OUT2
````

```{r WriteOut}
#write.table(anc.OUT1, "Predictions+Stats+ConfIntervals.csv", sep=";", append=TRUE)
#write.table(anc.OUT2, "Predictions+Stats+ConfIntervals.csv", sep=";", append=TRUE)
```

```{r PredictionPlot2}
  endo.bm.td = expand.grid(age_class=unique(data$age_class),
                     scl.pH = seq(min(data$scl.pH),max(data$scl.pH),by=0.2),
                     area = 1)

  endo.bm.pred <- cbind(  endo.bm.td, predict(	endo.bm.best, newdata = 	endo.bm.td, interval = "confidence")) 

ggplot(	endo.bm.pred, aes(x = scl.pH, y = exp(fit), ymin = exp(lwr), ymax = exp(upr))) + 
  geom_errorbar(position = position_dodge(1)) + 
  geom_point(position = position_dodge(1)) +
  facet_grid(.~age_class) +
  mytheme
```

# Results

```{r CreateREADME.md, }
# knit(input="README.Rmd", output="README.md")
```



