---
title: "Anecic Abundance"
author:
  - name: Quentin Schorpp
    email: qurntin.schorpp@ti.bund.de
date: "`r Sys.Date()`"
output: 
      html_document:
          toc: true
          toc_depth: 3
          number_sections: true
          fig_caption: true
          theme: cerulean
          highlight: haddock
          
---
```{r SetGlobalOptions, include=FALSE}
knitr::opts_chunk$set(fig.align="center",
                      message=FALSE,
                      error=FALSE,
                      warning=FALSE,
                      echo=FALSE)
knitr::opts_knit$set(root.dir = 'D://Quentin_Schorpp//Arbeitsprozess//git_repositories//F2_Earthworms')

options(scipen=1, digits=2)
set.seed(1234)
```


# Introduction  

I investigated how earthworm communities change in perennial energy farming systems. To assess the development in the long term, I chose an approach called Space-for-Time Substitution (Pickett, 1989). Instead of sampling sites over a long time period, I studied different aged sites. The sites were cropped either with the perennial bioenergy plant cup-plant (*Silphium perfoliatum*) or with annual Maize (*Zea mays*). The Main question of this investigation is wether older cup plant fields possess higher structural and functional biodiversity than younger ones and to compare cup-plant to maize. To distinguish the age-effect from environmental influences, several parameters regarding soil properties, climate and Management were collected.
Sampling took place 3-4 times in two consecutive years. In total 18 sites were surveyed, comprising 4 cup-plant age classes and maize. The sites were located at seven different locations, comprising fields from research and business. Each cup-plant age-class had three replicates, whereas maize plants were switched in the second year of sampling due to crop rotation and gained six replicates. To sample earthworms 4 pits with 0.25 m² were digged up to 20 cm in depth and both handsorting of soil and chemical extractioon with AITC took place. These 4 pits were randomly chosen at the field, hence they're nested in field.
9 earthworm species from 3 different ecological groups were identified and weighed.

Here I present the modeling of the abundance data of earthworms from the ecological group of anecic earthworms. 

# Materials and methods  

I use R with RStudio to perform all analyses. For creating reports i use knitr with RStudio.

## Modeling procedure:

* I start with a global model, with a full set of parameters. 
  
    + The choice of parameters took place according to preknowledge  about the biology of earthworms. 
    + Management parameters, are not included up to now. 

* Since Abundance is count data, i choose poisson distribution in the first place, and switch to negative binomial if overdipersion can be detected.

* As random effects i chose field. I could have chosen sampling campaign, too and analyse the model with crossed random effects structure, but I'm also interested in change over the time of sampling.

* I included an offset "area", since i want to present the data per m² not per 0.25 m².

* several global models were formulated if there are different combinations of parameters, meaningful for analysis, but not able to analyse at once, i.e. because of  collinearity.

* Then I use Multimodel Inference/Averaging to find the model with the lowest AIC and determine it as the best model.

* Model validation is performed by 

    + plotting residuals against all covariates in the model as well as against all covariates not in the model
    + plotting a histogram of the residual
    + checking for overdispersion

* after having found a model with good model fit, i present the data for the means of environmental covariates and plot barplots for the age classes.

* errorbars are calculated, using the inverse link function with confidence intervals.

* Post-Hoc Multicomparisons are performed using Tukey Test with the glht() function.

# Analysis

```{r LoadData, results='hide', fig.show='hide'}
# Load Data
#setwd("D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/")
source("Data/GatherSource/Gather1.R") # Includes Correction for Watercontent
source("Data/GatherSource/Gather2.R")
source("Data/GatherSource/Gather3.R") # no adult endogeics separated
source("Data/GatherSource/CNRatio.R") # only for 1st order
```

```{r RequiredPackages}
source("Analysis/F2_EW_RequiredPackages.R")
source("Data/GatherSource/Functions/coldiss.R")
source("Data/GatherSource/Functions/panelutils.R")
source("Data/GatherSource/ggplotTheme.R")

lo <- par(mfrow=c(1,1))

overdisp_fun <- function(model) {
  ## number of variance parameters in 
  ##   an n-by-n variance-covariance matrix
  vpars <- function(m) {
    nrow(m)*(nrow(m)+1)/2
  }
  model.df <- sum(sapply(VarCorr(model),vpars))+length(fixef(model))
  rdf <- nrow(model.frame(model))-model.df
  rp <- residuals(model,type="pearson")
  Pearson.chisq <- sum(rp^2)
  prat <- Pearson.chisq/rdf
  pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
  c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

```


### Data Pre-Processing  
#### Standardize Explanatory Variables  

```{r DataProcessing,}

## Create variable d(delta)samcam as the time difference of each sampling date to the first date, that each field was sampled; first date is Zero for all fields. ####
data$dsamcam <- with(data, round(c(rep(0,60),date[61:120]-date[1:60],date[121:180]-date[1:60]),0))

## Standardize all explanatory variables ####

# extract all numerical variables
std.var <- data[sapply(data,is.numeric)]
# exclude all response variables
std.var <- std.var[,-c(7:30,49:58)]

# standardize, each column is divided by their standard deviation, and the mean of each column is subtracted 
std.var <- as.data.frame(scale(std.var, center=TRUE, scale=TRUE))
colnames(std.var) <- paste("scl", colnames(std.var), sep = ".") # rename columns

data$area <- rep(0.25,180)

data <- cbind(data, std.var)


# Ordered or not??
# Since i found it too difficult to handle, i used non-ordered factors
data$age_class <- factor(data$age_class, ordered=TRUE)
data$samcam <- factor(data$samcam, ordered=FALSE)

#----
```


### Anecic Abundance
********************

#### Assess variability in random effects
```{r BoxPlots,echo=FALSE}

# Variability within sites
boxplot(anc~field.ID, data, col="grey", main="Variability within sites", xlab="sites orderd in decreasing age", ylab="anecic earthworm abundance")

# variability within sampling campaigns
boxplot(anc~samcam, data,col="grey", main="Variability within sampling campaigns", xlab="seasons:\n autumn2012, spring2013, autumn2013", ylab="anecic earthworm abundance")

# variability within sites and sampling campaigns
boxplot(anc~field.ID+samcam, data, las=2, col="grey", main="Variability within sites \n differing at sampling campaigns", xlab="field+seasons", ylab="anecic earthworm abundance")
```
  

#### GLMM using glmmADMB


##### Model Formulation

Find the best model using function dredge(). Begin with three different global models.
```{r GLMM_anc, cache=TRUE, echo=TRUE, eval=FALSE}
# Global Model

# str(data)

# Model Formulation
anc.glob1 <- glmmadmb(anc ~ age_class*samcam + scl.mc + scl.pH*scl.cn  +I(scl.hum1^2) + (1|field.ID) + offset(log(area)),data=data,family="poisson")
anc.glob2 <- glmmadmb(anc ~ age_class*samcam + I(scl.mc^2) + scl.pH*scl.cn  +I(scl.hum1^2) + (1|field.ID) + offset(log(area)),data=data,family="poisson")
anc.glob3 <- glmmadmb(anc ~ age_class*samcam + I(scl.ats1^2) + scl.pH*scl.cn  +I(scl.hum1^2) + (1|field.ID) + offset(log(area)),data=data,family="poisson")
# offsset is used due to the personal advice by T. Onkelinx:
# You better use an offset if you want to express the model in terms of m². Just add offset(log(0.25)) to the model. 

# summary(anc.glob3)

# Overdispersion
E1 <- resid(anc.glob3, type="pearson")
N <- nrow(data)
p <- length(coef(anc.glob3)) # The Plus "+1" used for determining the number of parameters(p) is due to the "k" of negative 
                              # binomial distribution
Dispersion <- sum(E1^2)/(N-p)
Dispersion

## Multimodel averaging ####
#anc.dredge1 <- dredge(anc.glob1)
#anc.dredge2 <- dredge(anc.glob2)
#anc.dredge3 <- dredge(anc.glob3)
#head(anc.dredge3,10)
#head(anc.dredge2,10)
#head(anc.dredge,10)
#####
````

Fit the best model
```{r BestModel, cache=TRUE, echo=TRUE}
# The best model with both glmer() and glmmadmb():
# anc.best <- glmer(anc ~ age_class*samcam + I(scl.ats1^2) + (1|field.ID) + offset(log(area)) ,data=data,family="poisson")
anc.best <- glmmadmb(anc ~ age_class*samcam + I(scl.ats1^2) +  (1|field.ID) + offset(log(area)) ,data=data,family="poisson")
````

**The best model includes an Interaction term!!!**


Summary output
```{r BestModel_Summaries, echo=TRUE}
# summary
summary(anc.best)

# Anova
summary(aov(anc.best))
````

A former "best model" with humidity as significant term and age_class as ordered factor, found using dredge was
anc.best <- glmmadmb(anc ~ age_class + scl.dsamcam + I(scl.ats1^2) +I(scl.hum1^2) + (1|field.ID) + offset(log(area)) ,data=data,family="poisson")

##### Model Validation
```{r ModelValidation1_anc, echo=TRUE}
# Confidence Intervals
confint(anc.best)
coefplot2(anc.best)


## Model validation ####
E1 <- resid(anc.best, type="pearson")
F1 <- fitted(anc.best, type="response")

# Dataset with all variables IN the model
env1 <- cbind(E1, F1, data[,c("anc", "age_class", "samcam", "ats1")]) # should "ats1" be squared?

# covariates NOT in the model
env0 <- cbind(E1, F1,data[,c("mc", "pH", "cn", "clay", "ats2", "hum1","dgO")])

# plot residual versus all covariates in the model
par(mfrow=c(2,3),
    mar=c(3.8,4,1,3))
for (i in 1:length(env1)) {
  scatter.smooth(env1[,i],env1[,1], xlab=colnames(env1)[i])
  abline(h=0, lty=2, col="red")
}
par(mfrow=c(1,1))

# plot residual versus all covariates NOT in the model
par(mfrow=c(3,3),
    mar=c(3.8,4,1,3))
for (i in 1:length(env0)) {
  scatter.smooth(env0[,i],env0[,1], xlab=colnames(env0)[i])
  abline(h=0, lty=2, col="red")
}
par(mfrow=c(1,1))

# Overdispersion
N <- nrow(data)
p <- length(coef(anc.best)) # +1 in case of negbin
Dispersion <- sum(E1^2)/(N-p)
Dispersion
# glmer has less dispersion

overdisp_fun(anc.best)
#####

# Deviance
logLik(anc.best) 

# Fraction of explained variation in the response variable
100*((anc.best$null.deviance-anc.best$deviance)/anc.best$null.deviance) # failed


# Plots of Predicted Values
par(mfrow=c(2,2))
plot(data$age_class,predict(anc.best, type="response"))
plot(data$samcam,predict(anc.best, type="response"))
plot(data$ats1^2,predict(anc.best, type="response"))
plot(data$hum1^2,predict(anc.best, type="response"))

# Plots of fitted Values
par(mfrow=c(2,2))
plot(data$age_class,F1)
plot(data$samcam,F1)
plot(data$ats1^2,F1)

par(lo)
```

Model validation 2, what is the meaning of that all?
```{r ModelValidation2_anc,fig.width=8, fig.height=7, echo=TRUE}
# Model Validation 

E <- resid(anc.best, type="pearson")

par(mfrow=c(2,2))
par(mar=c(5,5,2,3))
plot(E~fitted(anc.best), ylab="Pearson Residuals", xlab="Fitted values") 
abline(h=0, lty=2)
hist(E, main = "", breaks = 20, cex.lab = 1.5, xlab = "Pearson Residuals")
plot(E ~ data$age_class, ylab="Pearson Residuals", xlab="Age Class")
abline(h=0, lty=2)
plot(E ~ data$ats1, ylab="Pearson Residuals", xlab="Average Temperature (30d, 5cm)")
abline(h=0, lty=2)
# Cooks Distances?

````

##### Post-Hoc Multicomparisons
```{r PostHocMulticomparisons_anc, echo=TRUE, cache=TRUE}
# Model with the interaction term
data$ia.acl.smc <- interaction(data$age_class, data$samcam)
anc.tukey <- glmmadmb(anc ~ ia.acl.smc + I(scl.ats1^2) + (1|field.ID) + offset(log(area)) ,data=data,family="poisson")
# summary(anc.tukey)


# Pairwise comparisons
anc.pairwise <- glht(anc.best, mcp(age_class = "Tukey"))
anc.pw.ci <- confint(anc.pairwise)

# Confidence intervals including 0
anc.pw.sig <- which(anc.pw.ci$confint[,2]>0)
data.frame(names(anc.pw.sig))


# Plot Errorbars
ggplot(anc.pw.ci, aes(y = lhs, x = exp(estimate), xmin = exp(lwr), xmax = exp(upr))) + 
  geom_errorbarh() + 
  geom_point() + 
  geom_vline(xintercept = 1) +
  mytheme

# plot confidence intervals
par(mar=c(2,15,2,2))
plot(anc.pw.ci) # only maize is significantly different from all SIlphie fields. Within SIlphie there are no differences

```

##### Extract Predictions and plot predictions with error bars
```{r ExtractPredicitons, echo=TRUE}

# create test data set with all covariates IN the model
# to predict for age_class only, take the mean of all continous covariates
anc.td = expand.grid(age_class=unique(data$age_class),
                     samcam = unique(data$samcam),               
                     scl.ats1 = mean(data$scl.ats1),
                     area = 1)

# calculate confidence intervals for predictions from test dataset
anc.pred <- cbind(anc.td, predict(anc.best, newdata = anc.td, interval = "confidence")) 

# Rename samcam for facetting
anc.pred$samcam2 <- plyr::revalue(anc.td$samcam,c("1" ="autumn 2012",  "2" ="spring 2013", "3"="autumn 2013"))

# plot predictions with error bars // confidence intervals???
ggplot(anc.pred, aes(x = age_class, y = exp(fit), ymin = exp(lwr), ymax = exp(upr))) + 
  geom_bar(stat="identity",position = position_dodge(1), col="454545", fill="grey") +
  geom_errorbar(position = position_dodge(1),col="black",width=0.15, size=0.15) + 
  facet_grid(.~samcam2) +
  geom_hline(xintercept = 1) +
  ylab("Abundance Ind./m²") +
  xlab("Age Class") +
  scale_x_discrete(labels=c("Cm", "Sp_Y", "Sp_I1", "Sp_I2", "Sp_O")) +
  mytheme
````

Coefficients and Statistics
```{r, echo=TRUE}
# Claculate a whole lot of coefficients and statistics
anc.pred <- within(anc.pred, {
  AIC <- AIC(anc.best)
  Rrandom <- summary(lm(fitted(anc.best)~ data$anc))$adj.r.squared
  Rsquared <- summary(lm(predict(anc.best, type="response")~ data$anc))$adj.r.squared  
})
head(anc.pred)

anc.stat = round(summary(anc.best)$coefficients[, c(3,4)],4)
anc.coef = coeftab(anc.best)

anc.env <- glmmadmb(anc ~ samcam + I(ats1^2) + (1|field.ID) + offset(log(area)),data=data,family="poisson")
anc.pvalue <- anova(anc.env, anc.best)$"Pr(>Chi)"

anc.OUT1 = anc.pred
anc.OUT2 = cbind(anc.stat,anc.coef,LogLikP=anc.pvalue[2],fixef=fixef(anc.best))
anc.OUT1
anc.OUT2
````

```{r WriteOut}
#write.table(anc.OUT1, "Predictions+Stats+ConfIntervals.csv", sep=";", append=TRUE)
#write.table(anc.OUT2, "Predictions+Stats+ConfIntervals.csv", sep=";", append=TRUE)
```

##### Prediction plots for average temperature!
```{r PredictionPlot2}
anc.td = expand.grid(age_class=unique(data$age_class),
                     samcam = unique(data$samcam),               
                     scl.ats1 = seq(min(data$scl.ats1),max(data$scl.ats1), by=0.2),
                     area = 1)

anc.pred <- cbind(anc.td, predict(anc.best, newdata = anc.td, interval = "confidence")) 

ggplot(anc.pred, aes(x = scl.ats1, y = exp(fit), ymin = exp(lwr), ymax = exp(upr),col=samcam)) + 
  geom_errorbar(position = position_dodge(1)) + 
  geom_point(position = position_dodge(1)) +
  facet_grid(.~age_class) +
  mytheme
```


# Results

```{r CreateREADME.md, }
# knit(input="README.Rmd", output="README.md")
```



