---
title: "Anecic Abundance"
author:
  - name: Quentin Schorpp
    email: qurntin.schorpp@ti.bund.de
date: "`r Sys.Date()`"
output: 
      html_document:
          toc: true
          toc_depth: 2
          number_sections: true
          fig_caption: true
          theme: cerulean
          highlight: haddock
          
---
```{r SetGlobalOptions, include=FALSE}
knitr::opts_chunk$set(fig.align="center",
                      message=FALSE,
                      error=FALSE,
                      warning=FALSE,
                      echo=FALSE)

options(scipen=1, digits=2)
```


# Introduction  

I investigated how earthworm communities change in perennial energy farming systems. To assess the development in the long term, I chose an approach called Space-for-Time Substitution (Pickett, 1989). Instead of sampling sites over a long time period, I studied different aged sites. The sites were cropped either with the perennial bioenergy plant cup-plant (*Silphium perfoliatum*) or with annual Maize (*Zea mays*). The Main question of this investigation is wether older cup plant fields possess higher structural and functional biodiversity than younger ones and to compare cup-plant to maize. To distinguish the age-effect from environmental influences, several parameters regarding soil properties, climate and Management were collected.
Sampling took place 3-4 times in two consecutive years. In total 18 sites were surveyed, comprising 4 cup-plant age classes and maize. The sites were located at seven different locations, comprising fields from research and business. Each cup-plant age-class had three replicates, whereas maize plants were switched in the second year of sampling due to crop rotation and gained six replicates. To sample earthworms 4 pits with 0.25 m² were digged up to 20 cm in depth and both handsorting of soil and chemical extractioon with AITC took place. These 4 pits were randomly chosen at the field, hence they're nested in field.
9 earthworm species from 3 different ecological groups were identified and weighed.

Here I present the modeling of the abundance data of earthworms from the ecological group of anecic earthworms. 

## Modeling procedure:

* I start with a global model, with a full set of parameters. 
  
    + The choice of parameters took place according to preknowledge  about the biology of earthworms. 
    + Management parameters, are not included up to now. 

* Since Abundance is count data, i choose poisson distribution in the first place, and switch to negative binomial if overdipersion can be detected.

* As random effects i chose field. I could have chosen sampling campaign, too and analyse the model with crossed random effects structure, but I'm also interested in change over the time of sampling.

* I included an offset "area", since i want to present the data per m² not per 0.25 m².

* several global models were formulated if there are different combinations of parameters, meaningful for analysis, but not able to analyse at once, i.e. because of  collinearity.

* Then I use Multimodel Inference/Averaging to find the model with the lowest AIC and determine it as the best model.

* Model validation is performed by 

    + plotting residuals against all covariates in the model as well as against all covariates not in the model
    + plotting a histogram of the residual
    + checking for overdispersion

* after having found a model with good model fit, i present the data for the means of environmental covariates and plot barplots for the age classes.

* errorbars are calculated, using the inverse link function with confidence intervals.

* Post-Hoc Multicomparisons are performed using Tukey Test with the glht() function.



# Materials and methods  

# Analysis

# Results

```{r CreateREADME.md, }
# knit(input="README.Rmd", output="README.md")
```



