---
title: "F2_Earthworms"
author:
  - name: FirstName LastName
    email: name@company.com
date: "`r Sys.Date()`"
e-mail: quentin.schorpp@ti.bund.de
output: 
      html_document:
          toc: true
          toc_depth: 3
          number_sections: true
          fig_caption: true
          theme: cerulean
          highlight: haddock
---
```{r SetGlobalOptions, include=FALSE}
knitr::opts_chunk$set(fig.align="center",
                      message=FALSE,
                      error=FALSE,
                      warning=FALSE,
                      echo=FALSE)
knitr::opts_knit$set(root.dir = 'D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/')

options(scipen=1, digits=2)
set.seed(1234)
```

```{r LoadData, results='hide'}
# Load Data
#setwd("D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/")
source("Data/GatherSource/Gather1.R")
source("Data/GatherSource/Gather2.R")
source("Data/GatherSource/Gather3.R")
```

```{r RequiredPackages}
source("Analysis/F2_EW_RequiredPackages.R")
source("Data/GatherSource/coldiss.R")
source("Data/GatherSource/panelutils.R")
source("Data/GatherSource/ggplotTheme.R")
```


# Introduction  

# Materials and methods  

# Analysis

*******************************************************************************************
### Variable Descriptions  

Earthworm species:             || Environmental:
:-------------------------------------|---|:-----------------------------------------------------
ACA - *Aporrectodea caliginosa*      || crop [Silphie, Maize]
ACH - *Allolobophora chlorotica*     || soiltype [Ut4, Lt, etc.]
ALO - *Apporectodea longa*           || pH
ARO - *Apporectodea rosea*           || CN [ratio]
END - Endogeic, *spec.*              || moist [volumetric %]
LCA - *Lumbricus castaneus*          || sand, silt, clay content [%]
LRU - *Lumbricus rubellus*           || ats = average soil temperature in 5cm depth [?C] 
LTR - *Lumbricus terrestris*         || ata = average air temperature in 2 m height [?C]
LUM - *Lumbricus spec.*              || atb = average air temperature in 20 cm height [?C]
OCY - *Octolasion cyaneum*           || hum = air humidity [%]
OLA - *Octolasion lacteum*           || prec = precipitation [mm]
UNK - Unknown                        || rad = radiation

___________________________________________________________________________________________________

**Spatial:**                         || **Temporal:** 
:-------------------------------------|---|:------------------------------------------------------
field.ID [Number, representing field] || season [autumn, spring]
field [Field code, DO_2004,etc.]      || age class [old, intermediate 2, intermediate 1, young, maize]
hole [hole; 1-4]                      || year [year of sampling; 2012,2013]
location [Name of county]             || date [da of sampling]
dgO [degree East, longitude]          || year established [2004,-05, -07, -09, 11]
dgN [degree North, latitude]          || age [age of field; y]


**Functional groups:** || 
---|---|---
anc - anecic  ||
endo - endogeic  ||
epi - epigeic||

___________________________________________________________________________________________________

> * Environmental data: 1 = Average of 30 days before sampling; 2 = Average of day of sampling  
  * _BM = Biomass values for Earthworm species in [g]  

___________________________________________________________________________________________________

## Exploratory Data Analysis (EDA)

### Density Plots
```{r DensityPlots, fig.height=6, fig.width=7}
par(mfrow=c(2,2))
plot(density(data$N), main="density-plot of total abundance")
plot(density(data$anc), main="density-plot of anecic abundance")
plot(density(data$epi), main="density-plot of epigeic abundance")
plot(density(data$endo), main="density-plot of endogeic abundance") # Outliers > 80 ?
```

### Scatterplots for all pairs of environmental variables
```{r EnvironmentalPairplots,message=FALSE, warning=FALSE, echo=FALSE}
# Choose interesting environmental parameters
env.chosen <- data[,c("anc", "mc", "pH", "cn", "clay", "ats1", "ats2", "hum1", "dgN", "dgO")]
env.chosen <- data[,c("anc", "ats1", "prec1","ats2", "hum2", "ata2", "prec2")]
env.chosen <- data[,c("anc", "age_class", "samcam")]
env.chosen <- data[,c("anc", "age_class", "dgO")]
env.chosen <- data[,c("anc", "age_class", "location")]

# Bivariate plots with histograms on the diagonal and smooth fitted curves
op <- par(mfrow=c(1,1), pty="s")
pairs(env.chosen, lower.panel=panel.smooth, upper.panel=panel.cor, method="pearson", diag.panel=panel.hist, main= "Bivariate Plots with Histograms and smooth curves\n(pearson)")
par(op)

# correlated with anc:
# ats2, pH*cn, age_class, samcam, dgO
```

### Figures of real data for functional groups
```{r RealDataFigures}

# create subset of real data with age_class, abundance (plus N) and biomass
data.raw.fig <- data[,c(which(colnames(data)=="age_class"),                      
                        which(colnames(data)=="anc"):which(colnames(data)=="epi"),
                        which(colnames(data)=="N"),
                        which(colnames(data)=="anc.bm"):which(colnames(data)=="epi.bm"),
                        which(colnames(data)=="N.bm"))]


# melt frame for use in ggplot
data.rf.melted <- melt(data.raw.fig, id.vars=c(1,6:9), variable="sfg", value.name="abundance")
data.rf.melted <- data.rf.melted[,c(1,6,7)]
data.rf.melted <- ddply(data.rf.melted, c("age_class", "sfg"), summarise, 
                          abc.n=length(abundance),
                          abc.sum=sum(abundance),
                          abc.mean=mean(abundance),
                          abc.sd=sd(abundance),
                          abc.se=abc.sd/sqrt(abc.n)) 

data.rf.melted2 <- melt(data.raw.fig, id.vars=c(1,2:5), variable="sfg.bm", value.name="biomass")
data.rf.melted2 <- data.rf.melted2[,c(1,6,7)]
data.rf.melted2 <- ddply(data.rf.melted2, c("age_class", "sfg.bm"), summarise, 
                          bm.n=length(biomass),
                          bm.sum=sum(biomass),
                          bm.mean=mean(biomass),
                          bm.sd=sd(biomass),
                          bm.se=bm.sd/sqrt(bm.n)) 

data.rf <- cbind(data.rf.melted,data.rf.melted2[,c(3:7)])

# Barplots
ggplot(data.rf, aes(x=age_class, y=4*abc.mean, fill=sfg)) +  
  geom_bar(stat="identity", position="dodge", colour="#454545", size=0.15) + 
  geom_errorbar(aes(ymin=4*abc.mean-4*abc.se, ymax=4*abc.mean+4*abc.se), position=position_dodge(0.9),width=0.15, size=0.15) +
  ggtitle("Data 1st Order") +
  xlab("Age Class") + 
  ylab("Abundance") +
  ylim(-10,max(data.rf$abc.mean+data.rf$abc.se)) +
  labs(fill="Functional Group") +
  scale_fill_grey(labels=c("anecic","endogeic", "epigeic", "total")) +
  scale_y_continuous(breaks=pretty_breaks(n=10)) +
  scale_x_discrete(labels=c("Cm", "Sp_Y", "Sp_I1", "Sp_I2", "Sp_O")) +  
  mytheme +
  theme(axis.text.x =element_text(angle=30, hjust=1, vjust=1),
        legend.position=c(0.18,0.68))
```

Version 2 with "2nd Order" data
```{r}
# create subset of real data2 with age_class, abundance (plus N) and biomass
data2.raw.fig <- data2[,c(which(colnames(data2)=="age_class"),                      
                        which(colnames(data2)=="anc"):which(colnames(data2)=="epi"),
                        which(colnames(data2)=="N"),
                        which(colnames(data2)=="anc.bm"):which(colnames(data2)=="epi.bm"),
                        which(colnames(data2)=="N.bm"))]


# melt frame for use in ggplot
data2.rf.melted <- melt(data2.raw.fig, id.vars=c(1,6:9), variable="sfg", value.name="abundance")
data2.rf.melted <- data2.rf.melted[,c(1,6,7)]
data2.rf.melted <- ddply(data2.rf.melted, c("age_class","sfg"), summarise, 
                          abc.n=length(abundance),
                          abc.sum=sum(abundance),
                          abc.mean=mean(abundance),
                          abc.sd=sd(abundance),
                          abc.se=abc.sd/sqrt(abc.n)) 

data2.rf.melted2 <- melt(data2.raw.fig, id.vars=c(1,2:5), variable="sfg.bm", value.name="biomass")
data2.rf.melted2 <- data2.rf.melted2[,c(1,6,7)]
data2.rf.melted2 <- ddply(data2.rf.melted2, c("age_class","sfg.bm"), summarise, 
                          bm.n=length(biomass),
                          bm.sum=sum(biomass),
                          bm.mean=mean(biomass),
                          bm.sd=sd(biomass),
                          bm.se=bm.sd/sqrt(bm.n)) 

data2.rf <- cbind(data2.rf.melted,data2.rf.melted2[,c(2:7)])

# Barplots
ggplot(data2.rf, aes(x=age_class, y=abc.mean, fill=sfg)) +  
  geom_bar(stat="identity", position="dodge", colour="#454545", size=0.15) + 
  geom_errorbar(aes(ymin=abc.mean-abc.se, ymax=abc.mean+abc.se), position=position_dodge(0.9),width=0.15, size=0.15) +
  ggtitle("Data 2nd Order") +
  xlab("Age Class") + 
  ylab("Abundance") +
  ylim(-10,max(data2.rf$abc.mean+data2.rf$abc.se)) +
  labs(fill="Functional Group") +
  scale_fill_grey(labels=c("anecic","endogeic", "epigeic", "total")) +
  scale_y_continuous(breaks=pretty_breaks(n=10)) +
  scale_x_discrete(labels=c("Cm", "Sp_Y", "Sp_I1", "Sp_I2", "Sp_O")) +  
  mytheme +
  theme(axis.text.x =element_text(angle=30, hjust=1, vjust=1),
        legend.position=c(0.18,0.68))

```

Version 3 with "3rd Order" data
```{r}
# create subset of real data2 with age_class, abundance (plus N) and biomass
data3.raw.fig <- data3[,c(which(colnames(data3)=="age_class"),                      
                        which(colnames(data3)=="anc"):which(colnames(data3)=="epi"),
                        which(colnames(data3)=="N"),
                        which(colnames(data3)=="anc.bm"):which(colnames(data3)=="epi.bm"),
                        which(colnames(data3)=="N.bm"))]


# melt frame for use in ggplot
data3.rf.melted <- melt(data3.raw.fig, id.vars=c(1,6:9), variable="sfg", value.name="abundance")
data3.rf.melted <- data3.rf.melted[,c(1,6,7)]
data3.rf.melted <- ddply(data3.rf.melted, c("age_class","sfg"), summarise, 
                          abc.n=length(abundance),
                          abc.sum=sum(abundance),
                          abc.mean=mean(abundance),
                          abc.sd=sd(abundance),
                          abc.se=abc.sd/sqrt(abc.n)) 

data3.rf.melted2 <- melt(data3.raw.fig, id.vars=c(1,2:5), variable="sfg.bm", value.name="biomass")
data3.rf.melted2 <- data3.rf.melted2[,c(1,6,7)]
data3.rf.melted2 <- ddply(data3.rf.melted2, c("age_class","sfg.bm"), summarise, 
                          bm.n=length(biomass),
                          bm.sum=sum(biomass),
                          bm.mean=mean(biomass),
                          bm.sd=sd(biomass),
                          bm.se=bm.sd/sqrt(bm.n)) 

data3.rf <- cbind(data3.rf.melted,data3.rf.melted2[,c(2:7)])

## Barplots ####

# Abundance
ggplot(data3.rf, aes(x=age_class, y=abc.mean, fill=sfg)) +  
  geom_bar(stat="identity", position="dodge", colour="#454545", size=0.15) + 
  geom_errorbar(aes(ymin=abc.mean-abc.se, ymax=abc.mean+abc.se), position=position_dodge(0.9),width=0.15, size=0.15) +
  ggtitle("Data 3rd Order") +
  xlab("Age Class") + 
  ylab("Abundance") +
  ylim(-10,max(data3.rf$abc.mean+data3.rf$abc.se)) +
  labs(fill="Functional Group") +
  scale_fill_grey(labels=c("anecic","endogeic", "epigeic", "total")) +
  scale_y_continuous(breaks=pretty_breaks(n=10)) +
  scale_x_discrete(labels=c("Cm", "Sp_Y", "Sp_I1", "Sp_I2", "Sp_O")) +  
  mytheme +
  theme(axis.text.x =element_text(angle=30, hjust=1, vjust=1),
        legend.position=c(0.18,0.68))


# ggsave("Figure1.pdf", width=9, height=7, units="cm", useDingbats=FALSE)
# ggsave("Figure1.png", width=9, height=7, units="cm")

# Biomass
ggplot(data3.rf, aes(x=age_class, y=bm.mean, fill=sfg)) +  
  geom_bar(stat="identity", position="dodge", colour="#454545", size=0.15) + 
  geom_errorbar(aes(ymin=bm.mean-bm.se, ymax=bm.mean+bm.se), position=position_dodge(0.9),width=0.15, size=0.15) +
  ggtitle("Data 3rd Order") +
  xlab("Age Class") + 
  ylab("Biomass") +
  ylim(-10,max(data3.rf$bm.mean+data3.rf$bm.se)) +
  labs(fill="Functional Group") +
  scale_fill_grey(labels=c("anecic","endogeic", "epigeic", "total")) +
  scale_y_continuous(breaks=pretty_breaks(n=10)) +
  scale_x_discrete(labels=c("Cm", "Sp_Y", "Sp_I1", "Sp_I2", "Sp_O")) +  
  mytheme +
  theme(axis.text.x =element_text(angle=30, hjust=1, vjust=1),
        legend.position=c(0.18,0.68)) 

# ggsave("Figure2.pdf", width=9, height=7, units="cm", useDingbats=FALSE)
# ggsave("Figure2.png", width=9, height=7, units="cm")
#####
```

> **Why are the error bars becoming larger and larger with the Order of the dataset?**

> One reason is that the standard error is the standard deviation divided by the squareroot of number of items.
The latter decreases with increasing order of dataset:  
The squareroot in 1st order is ``r sqrt(4*3*3)``  
The squareroot in 2nd order is ``r sqrt(3*3)``  
The squareroot in 3rd order is ``r sqrt(3)``   


```{r RemoveDatasets1, include=FALSE}
rm(data.raw.fig, data.rf, data.rf.melted, data.rf.melted2,
   data2.raw.fig, data2.rf, data2.rf.melted, data2.rf.melted2,
   data3.raw.fig, data3.rf, data3.rf.melted, data3.rf.melted2)
```

______________________________________________________________________________________________________________

## General linear Mixed Effects Models

### Data Pre-Processing  
#### Standardize Explanatory Variables  
```{r DataProcessing}

## Create variable d(delta)samcam as the time difference of each sampling date to the first date, that each field was sampled; first date is Zero for all fields. ####
data$dsamcam <- with(data, round(c(rep(0,60),date[61:120]-date[1:60],date[121:180]-date[1:60]),0))

## Standardize all explanatory variables ####

# extract all numerical variables
std.var <- data[sapply(data,is.numeric)]
# exclude all response variables
std.var <- std.var[,-c(7:30,49:58)]

# standardize, each column is divided by their standard deviation, and the mean of each column is subtracted 
std.var <- as.data.frame(scale(std.var, center=TRUE, scale=TRUE))
colnames(std.var) <- paste("scl", colnames(std.var), sep = ".") # rename columns

data <- cbind(data, std.var)
#----
```



### Anecic Abundance

```{r BoxPlots,echo=FALSE}

# Variability within sites
boxplot(anc~field.ID, data, col="grey", main="Variability within sites", xlab="sites orderd in decreasing age", ylab="anecic earthworm abundance")

# variability within sampling campaigns
boxplot(anc~samcam, data,col="grey", main="Variability within sampling campaigns", xlab="seasons:\n autumn2012, spring2013, autumn2013", ylab="anecic earthworm abundance")

# variability within sampling campaigns
boxplot(anc~field.ID+samcam, data, las=2, col="grey", main="Variability within sites \n differing at sampling campaigns", xlab="field+seasons", ylab="anecic earthworm abundance")
```
     
 

```{r GLMM, cache=TRUE}
M1 <- glmer(anc ~ age_class*scl.dsamcam + scl.mc*scl.cn + scl.ats2 + scl.dgO + (1|field.ID),data=data,family="poisson"); summary(M1)
M2 <- glmer(anc ~ age_class*scl.dsamcam + scl.dgO + (1|field.ID),data=data,family="poisson"); summary(M2)

M1 <- glmmadmb(anc ~ age_class*scl.dsamcam + scl.mc*scl.cn + scl.ats2 + scl.dgO + (1|field.ID),data=data,family="poisson"); summary(M1)
M2 <- glmmadmb(anc ~ age_class*scl.dsamcam + scl.dgO + (1|field.ID),data=data,family="poisson"); summary(M2)


AICtab(M1,M2)
anc.best <- M2
```



#### Model Validation
```{r ModelValidation,fig.width=8, fig.height=7}
# Model Validation 
confint(anc.best)
par(mfrow=c(2,2))
par(mar=c(5,5,2,3))
plot(resid(anc.best)~fitted(anc.best)) 
plot(fitted(anc.best) ~predict(anc.best, type="response")) 
plot(predict(anc.best) ~predict(anc.best, type="response")) 
plot(predict(anc.best, type="response") ~ data$anc)
````


#### Extract Predictions and Standard errors for predictions
```{r ExtractPredicitons}
# anc.1.mcmc = glmmadmb(anc ~ age_class + (1|field) + (1|samcam),data=ew.data,family="nbinom", zeroInflation=TRUE, mcmc = TRUE, mcmc.opts = mcmcControl(mcmc = 10000))

td.anc=expand.grid(age_class=levels(data$age_class),
              scl.dsamcam = mean(data$scl.dsamcam),
               scl.dgO = mean(data$scl.dgO))         
pred.anc <- cbind(td.anc,predict(anc.best, td.anc, type="link", se.fit=TRUE),anc.response=predict(anc.best, td.anc, type="response"))

pred.anc <- within(pred.anc, {
  AIC <- AIC(anc.best)
  Rrandom <- summary(lm(fitted(anc.best)~ data$anc))$adj.r.squared
  Rsquared <- summary(lm(predict(anc.best, type="response")~ data$anc))$adj.r.squared
  anc <- 4*exp(fit)
  LL <- 4*exp(fit - 1.96 * se.fit)   
  UL <- 4*exp(fit + 1.96 * se.fit)  
})
head(pred.anc)
range(data$anc)

stat.anc = round(summary(anc.best)$coefficients[, c(3,4)],4)
coef.anc = coeftab(anc.best)

anc.env <- glmmadmb(anc ~ scl.dsamcam + scl.dgO + (1|field.ID),data=data,family="poisson")
anc.pvalue <- anova(anc.env, anc.best)$"Pr(>Chi)"

pred.Stat.anc = cbind(stat.anc,coef.anc,LogLikP=anc.pvalue[2],fixef=fixef(anc.best))
````

```{r}
#write.table(pred.Stat.anc, "Predictions+Stats+ConfIntervals.csv", sep=";", append=TRUE)
#write.table(pred.anc, "Predictions+Stats+ConfIntervals.csv", sep=";", append=TRUE)
```

#### Prediciton plots
```{r AncPredictionPlot}
library(ggplot2)

ggplot(pred.anc, aes(x=age_class, y=anc)) + 
  geom_bar(stat="identity", width=0.5) +
  geom_line(aes(group=anc))+  
  geom_errorbar(aes(ymin=LL, ymax=UL), width=.1) +
  geom_line()


par(mfrow=c(1,1))
par(mar=c(3,3,3,3))
par(oma=c(1,1,1,1))
coefplot2(anc.best)    
```




# Results

### README
```{r CreateREADME.md, }
#setwd("D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/")
knitr::knit(input="README.Rmd", output="README.md") # Error: no such file or directory, when trying knitHTML
```



