---
title: "F2_Earthworms"
author:
  - name: FirstName LastName
    email: name@company.com
date: "`r Sys.Date()`"
e-mail: quentin.schorpp@ti.bund.de
output: 
      html_document:
          toc: true
          toc_depth: 2
          number_sections: true
          fig_caption: true
          theme: cerulean
          highlight: haddock
---
```{r SetGlobalOptions, include=FALSE}
knitr::opts_chunk$set(fig.align="center",
                      message=FALSE,
                      error=FALSE,
                      warning=FALSE,
                      echo=FALSE,
                      root.dir = 'D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/')

options(scipen=1, digits=2)
set.seed(1234)
```

```{r, results='hide'}
setwd("D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/")
source("Data/GatherSource/Gather1.R")
```


# Introduction  

# Materials and methods  

# Analysis

*******************************************************************************************
### Variable Descriptions  

Earthworm species:             || Environmental:
:-------------------------------------|---|:-----------------------------------------------------
ACA - *Aporrectodea caliginosa*      || crop [Silphie, Maize]
ACH - *Allolobophora chlorotica*     || soiltype [Ut4, Lt, etc.]
ALO - *Apporectodea longa*           || pH
ARO - *Apporectodea rosea*           || CN [ratio]
END - Endogeic, *spec.*              || moist [volumetric %]
LCA - *Lumbricus castaneus*          || sand, silt, clay content [%]
LRU - *Lumbricus rubellus*           || ats = average soil temperature in 5cm depth [?C] 
LTR - *Lumbricus terrestris*         || ata = average air temperature in 2 m height [?C]
LUM - *Lumbricus spec.*              || atb = average air temperature in 20 cm height [?C]
OCY - *Octolasion cyaneum*           || hum = air humidity [%]
OLA - *Octolasion lacteum*           || prec = precipitation [mm]
UNK - Unknown                        || rad = radiation

___________________________________________________________________________________________________

**Spatial:**                         || **Temporal:** 
:-------------------------------------|---|:------------------------------------------------------
field.ID [Number, representing field] || season [autumn, spring]
field [Field code, DO_2004,etc.]      || age class [old, intermediate 2, intermediate 1, young, maize]
hole [hole; 1-4]                      || year [year of sampling; 2012,2013]
location [Name of county]             || date [da of sampling]
dgO [degree East, longitude]          || year established [2004,-05, -07, -09, 11]
dgN [degree North, latitude]          || age [age of field; y]


**Functional groups:** || 
---|---|---
anc - anecic  ||
endo - endogeic  ||
epi - epigeic||

___________________________________________________________________________________________________

> * Environmental data: 1 = Average of 30 days before sampling; 2 = Average of day of sampling  
  * _BM = Biomass values for Earthworm species in [g]  

___________________________________________________________________________________________________

### Density Plots
```{r DensityPlots, fig.height=6, fig.width=7}
par(mfrow=c(2,2))
plot(density(data$N), main="density-plot of total abundance")
plot(density(data$anc), main="density-plot of anecic abundance")
plot(density(data$epi), main="density-plot of epigeic abundance")
plot(density(data$endo), main="density-plot of endogeic abundance") # Outliers > 80 ?
```


# Results

### README
```{r CreateREADME.md, }
setwd("D:/Quentin_Schorpp/Arbeitsprozess/git_repositories/F2_Earthworms/")
knitr::knit(input="README.Rmd", output="README.md") # Error: no such file or directory, when trying knitHTML
```



